---
title: |
  | **Probability \& Statistics I**
  | **Tentative Topics Schedule**
author: |
  | MTH 461 Section A
  | Fall 2024
  | University of Portland
bibliography: references.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(file.path("R","functions.R"))
```

```{r echo=FALSE}
create_footer()
```

<br>

> See Books \& Online Resources Lists for the readings \& practice materials.
> 
> The reading materials are not mandatory but it is encouraged.

<br>The "Reading" column in the table below contains page numbers (Pg.) or chapters (ch.) on which it refers to a label in the Books \& Online Resources List. For example "Pg. 1-5 [S]" refers to pages 1-5 of the first item in the list, which is the textbook titled "Probability, Statistics, and Data: A Fresh Approach Using R".

<br>

## **Topics and Materials**

| **Week** | **Day** | **Topic** | **Worksheet** | **Homework** | **Reading** |
|:---:|:----:|:---------------:|:------:|:-----:|:--------:|
| **1** | Tu 8/27 | Introduction and Orientation to <br>Probability \& Statistics | Review Set Theory \& Calculus | - | Syllabus |
|  | Th 8/29 | Basic Definition of Probability | Computing Probabilities and Random Sampling | - | Sect. 1.1-1.3 [B] |
| **2** | Tu 9/3 | Counting and Arranging | Introduce the Basics of Combinatorics | - | Sect. 1.4-1.5 [B] |
|  | Th 9/5 | General Definition of Probability | Understand the Properties of Probability | - | Sect. 1.6 [B] |
| **3** | Tu 9/10 | Conditional Probability | Think Conditionally | - | Sect. 2.1-2.2 [B] |
|  | Th 9/12 | Baye's Theorem, \&<br>The Law of Total Probability | Update Conditional Probabilities | Assigned: Homework 1 | Sect. 2.3-2.4 [B] |
| **4** | Tu 9/17 | Random Variables \&<br>Probability Functions | Describe a Random Variable | - | Sect. 3.1 [B] |
|  | Th 9/19 | Expectation \&<br>The Law of Large Numbers | Compute the Expected Value | - | Sect. 4.1-4.3 [B] |
| **5** | Tu 9/24 | Discrete Random Variables \&<br>Probability Mass Functions | Create Probability Mass Functions | - | Sect. 3.2-3.3 [B] |
|  | Th 9/26 | Discrete Probability Distributions | Understand Known Discrete Probability Distributions | Assigned: Homework 2 | Sect. 3.4-3.6 [B] |
| **6** | Tu 10/1 | _Cancelled_ | - | - | - |
|  | Th 10/3 | Functions of Random Variables \&<br>Variance of Discrete Random Variables | Determine Functions of Random Variables and their Probabilities \& Compute the Variance | - | Sect. 4.4-4.6 [B] |
| **7** | Tu 10/8 | _Review_ | Exam 1 Examples | - | Exam 1 Topics |
|  | Th 10/10 | **Midterm Exam** | - | - | - |
| **8** | Tu 10/15 | _Fall Vacation_ | - | - | - |
|  | Th 10/17 | _Fall Vacation_ | - | - | - |
| **9** | Tu 10/22 | Continuous Random Variables \&<br>Probability Density Functions | Create Probability Density functions | - | Sect. 5.1-5.3 [B] |
|  | Th 10/24 | Continuous Probability Distributions | Understand Known Continuous Probability Distributions | - | Sect. 5.4-5.6 [B] |
| **10** | Tu 10/29 | *Continued from 10/24* | - | - | - |
|  | Th 10/31 | Expectation and Variance of Continuous Random Variables | Compute the Expectation | - | Sect. 6.1-6.2 [B] |
| **11** | Tu 11/5 | *Continued from 10/31* | - | - | - |
|  | Th 11/7 | *Cancelled* | - | Assigned: Homework 3 | - |
| **12** | Tu 11/12 | Moment Generating Functions | Determine the Moments | - |  Sect. 6.3-6.5 [B] |
|  | Th 11/14 | Joint and Marginal Distributions | Compute the Marginals | - | Sect. 7.1-7.2 [B] |
| **13** | Tu 11/19 | *Continued from 11/14* | - | - | - |
|  | Th 11/21 | Covariance and Correlation | Determine the Covariance and Correlation | - | Sect. 7.3 [B] |
| **14** | Tu 11/26 | _Cancelled_ | - | - | - |
|  | Th 11/28 | _Thanksgiving Vacation_ | - | - | - |
| **15** | Tu 12/3 | Convolutions | TBA | - | Sect. 8.2 [B] |
|  | Th 12/5 | _Review_ | Exam 2 Examples | - | Exam 2 Topics |
| **16** | W 12/11 | **Final Exam Section A** | - | - | - |

<br>

## **Books \& Online Resources Lists**

```{r echo=FALSE}
refs <- bibtex::read.bib("references.bib")
```

[B] `r capture.output(print(refs["blitzstein2019introduction"]))`

*Amendment 7:* In relation to *Amendment 6*, the recommended textbooks are reduced from two to one, and the extra book resources were removed.

*Amendment 8:* The "Textbooks" subsection label is removed under the "Books \& Online Resources Lists".

<br>

## **References**

::: {#refs}
:::

<br>
