---
title: |
  | **Probability \& Statistics I**
  | **Tentative Topics Schedule**
author: |
  | MTH 461 Section A
  | Fall 2024
  | University of Portland
bibliography: references.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(file.path("R","functions.R"))
```

```{r echo=FALSE}
create_footer()
```

<br>

> See Books \& Online Resources Lists for the readings \& practice materials.
> 
> The reading materials are not mandatory but it is encouraged.

<br>The "Reading" column in the table below contains page numbers (Pg.) or chapters (ch.) on which it refers to a label in the Books \& Online Resources List. For example "Pg. 1-5 [S]" refers to pages 1-5 of the first item in the list, which is the textbook titled "Probability, Statistics, and Data: A Fresh Approach Using R".

<br>

## **Topics and Materials**

| **Week** | **Day** | **Topic** | **Worksheet** | **Homework** | **Reading** |
|:---:|:----:|:---------------:|:------:|:-----:|:--------:|
| **1** | Tu 8/27 | Introduction and Orientation to <br>Probability \& Statistics | Review Set Theory \& Calculus | - | Syllabus |
|  | Th 8/29 | Basic Definition of Probability | Computing Probabilities and Random Sampling | - | Sect. 1.1,1.2,\&1.3 [B] |
| **2** | Tu 9/3 | Counting and Arranging | Introduce the Basics of Combinatorics | - | Sect. 1.4 \& 1.5 [B], and<br>Sect. 2.4 [S] |
|  | Th 9/5 | General Definition of Probability | Understand the Properties of Probability | - | Sect. 1.6 [B], and<br>Sect. 2.1\&2.2 [S] |
| **3** | Tu 9/10 | Conditional Probability | Think Conditionally | - | Sect. 2.1 \& 2.2 [B], and<br>Sect. 2.3 [S] |
|  | Th 9/12 | Baye's Theorem, \&<br>The Law of Total Probability | Update Conditional Probabilities | Assigned: Homework 1 | Sect. 2.3 \& 2.4 [B] |
| **4** | Tu 9/17 | Random Variables \&<br>Probability Functions | Describe a Random Variable | - | Sect. 3.1 \& 3.2 [B] |
|  | Th 9/19 | Expectation \&<br>The Law of Large Numbers | Compute the Expected Value | - | TBA |
| **5** | Tu 9/24 | Discrete Random Variables \&<br>Probability Mass Functions | TBA | - | TBA |
|  | Th 9/26 | Bernoulli, Binomial, and Geometric Distributions | TBA | Assigned: Homework 2 | TBA |
| **6** | Tu 10/1 | Expectation and Variance of Discrete Random Variables | TBA | - | TBA |
|  | Th 10/3 | Continuous Random Variables \&<br>Probability Density Functions | TBA | - | TBA |
| **7** | Tu 10/8 | _Review_ | Exam 1 Examples | - | Exam 1 Topics |
|  | Th 10/10 | **Exam 1** | - | - | - |
| **8** | Tu 10/15 | _Fall Vacation_ | - | - | - |
|  | Th 10/17 | _Fall Vacation_ | - | - | - |
| **9** | Tu 10/22 | Uniform, Normal, and Exponential Distributions | TBA | - | TBA |
|  | Th 10/24 | Expectation and Variance of Continuous Random Variables | TBA | Assigned: Homework 4 | TBA |
| **10** | Tu 10/29 | Moment Generating Functions | TBA | - | TBA |
|  | Th 10/31 | Joint and Marginal Distributions | TBA | Assigned: Homework 5 | TBA |
| **11** | Tu 11/5 | Conditional Distributions | TBA | - | TBA |
|  | Th 11/7 | Conditional Expectation and Variance | TBA | Assigned: Homework 6 | TBA |
| **12** | Tu 11/12 | Maximum Likelihood Estimation | TBA | - | TBA |
|  | Th 11/14 | Markov Chains | TBA | - | TBA |
| **13** | Tu 11/19 | _Review_ | Exam 2 Examples | - | Exam 2 Topics |
|  | Th 11/21 | **Exam 2** | - | - | - |
| **14** | Tu 11/26 | Markov Chain Monte Carlo | TBA | - | TBA |
|  | Th 11/28 | _Thanksgiving Vacation_ | - | - | - |
| **15** | Tu 12/3 | Statistical Inference | TBA | - | TBA |
|  | Th 12/5 | Statistical Learning | TBA | - | TBA |
| **16** | Tu 12/11 | **Final Exam Section A** | - | - | - |

<br>

Along with textbooks [S] and [B], some of the course materials (contents of worksheets and homework) of each topic was taken from these following sources:

-   The elements of statistical learning: data mining, inference, and prediction by @hastie2009elements
-   An introduction to statistical learning with Applications in R by @james2013introduction

<br>

## **Books \& Online Resources Lists**

```{r echo=FALSE}
refs <- bibtex::read.bib("references.bib")
```

*Click on the link to access the resources.*

### Textbooks

[S] `r capture.output(print(refs["speegle2021probability"]))`

[B] `r capture.output(print(refs["blitzstein2019introduction"]))`

<br>

## **References**

::: {#refs}
:::

<br>
